### 2016-08-31 version0.0/paper2.tex
Adding an appendix about Bayesian fitting
Applying S.G.'s comment on changing Y title of Spectra plots
Update paper2.tex
grammatical corrections to paper2.tex

\chapter[Classifying High-Redshift Galaxies Spectra]{Clustering galaxy spectra at $0.5<z<1$: an unsupervised approach}
\label{ch: paper2}
\defcitealias{Hossein12}{T12}
\defcitealias{Kinney96}{K96}
\defcitealias{Noll09}{N09}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction}
\label{sec: intro_somz}
%General information about SEDs
Nearly all information we can obtain from a galaxy is encapsulated in the light it emits; every observable phenomenon in a galaxy leaves a footprint on its spectral energy distribution (SED).
We can determine some physical property of a galaxy by properly modeling various features observed in its SED, and the general shape of the SED can be used as an identifier of the morphological type.
Considering the main features of SEDs, galaxies can be categorized into two main groups: elliptical or spiral.
Each group has its own characteristic features and can in turn be divided into many sub-branches.

@@

Advances in techniques and detectors have resulted in more detailed observations of SEDs and also made classification of galaxies more complex.
Since no two galaxies, even with the same morphology, have exactly the same properties, classifying SEDs of galaxies using templates is very challenging.
To overcome this challenge, many fitting methods have been developed and used to find template matches for SEDs, with $\chi^2$ minimization being the most commonly used. 
Artificial neural networks, K-means clustering, and principal component analysis are some other methods used to cluster and classify spectral types of galaxies based on their SED \citep[e.g.][]{Allen13,Ordov14,Shi15}.

@@

%SOM
A Kohonen self-organizing map (also called self-organizing map, or SOM) is an unsupervised neural network for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
%SOM in Astronomy
The utilization of the SOM in astronomy dates back to the 1990s, with \citet[][]{Odewahn92}, \citet[][]{Hernandez94}, and \citet[][]{Murtagh95} among the first to use SOMs in their studies.
\citet{Geach12} used COSMOS data to demonstrate two of the main applications of SOMs: object classification and clustering, and photometric redshift estimation. The latter has been the subject of many other studies \citep[e.g.][]{Kind14a}.
From classifying quasars' spectra to star/galaxy classifications, from gamma-ray bursts clustering to classification of light curves, this method has proved to be useful in various fields of astronomy \citep[e.g.][]{Maehoenen95, Miller96, Andreon00, Balastegui01, Rajaniemi02, Brett04, Scaringi09}.

%SOM in Astronomy continued.
Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can recognize an astronomical object from instrumental errors, and then classify the object based on its spectrum.

%What \citetalias{Hossein12} did
~\citet[][hereafter \citetalias{Hossein12}]{Hossein12} classified spectral types of a sample of 142 galaxies using a supervised (Multi-layer feed-forward) neural network method, based on the spectral template presented by \citetalias{Kinney96}.
With the supervised method, they could only classify spectral types of 105 out of the 142 galaxies.
The spectra of the 37 galaxies from their samples could not be matched with any of those in the \citetalias{Kinney96} templates. 
In order to classify the remaining 37 galaxies, they combined spectrum of the \citetalias{Kinney96} templates.
\citetalias{Hossein12} also showed that there are tight correlations between physical properties of galaxies, and these correlations might be different for each type of galaxy.

@@

\section{Data}
\label{sec: data_highZ}
We use spectra templates from \citetalias{Kinney96} to train neural networks.
To test the trained networks, we use SED and physical properties of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12}.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in spectra clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney Spectral Model}
     \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{../image_paper2/k96.jpg}
        \caption[\citet{Kinney96} spectra template for 12 types of galaxies]{\citet{Kinney96} spectra template for 12 types of galaxies from \citet{Hossein12} paper (Fig. 1 in \citet{Hossein12}). The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and SC show spectra that belong to the quiescent galaxies. Starburst galaxies spectra are indicated with SB 1 to 6. Higher numbers represent more intrinsic colour extinction.}
        \label{fig: k96}
    \end{figure}

    \citetalias{Kinney96} used ultraviolet-optical spectra of 70 star forming and quiescent nearby galaxies to produce a set of templates that contained 12 types of spectra templates.
    These templates have been widely used in many studies to determine morphological type of galaxies or properties of specific types of galaxies~\citep[e.g.][]{Shakouri16, Paiano16, Laporte16, Holden16}.
    \citetalias{Kinney96} stated that these templates can also be used to classify the spectra of high-redshift galaxies.

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Method
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Method}
\label{sec: method_somz}
 \subsection{Self Organizing Maps}
 \label{sec: som_SOMZ}
  The SOM is a clustering method which reduces the dimensionality of data, while preserving topological features~\citep{Kohonen98}. 
 The results of the SOM are shown with a map of neurons, which their numbers are set by the user.
 Each neuron has a fixed position in the map and may contain one or more samples from the input data in a n-dimensional space,~\boldit{V} $\in \Re^n$.
 A weight vector,~\boldit{W} $\in \Re^n$, with the same dimension as the input data, is associated with each node and will be varied during the training process.
 The process of creating an SOM happens over a series of $N$ iterations.
 In each iteration, the algorithm calculates the Euclidean distance for each node $j$ as $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and finds a neuron with $D_{j_,\mathrm{min}}$. 
 This neuron is the winner node and is calling Best Matching Unit (BMU). 
 In each iteration (t), the weight vectors W(t) in the neighbourhood of the BMU ($r^t_{BMU}$) will change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
  \begin{equation}
            \label{equ: weight adj}
            W(t+1)=W(t)+L(t) \times R(t) \times(V(t)-W(t))
 \end{equation}
where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor with $L_0$ is the initial value of learning, which prevents the divergence of the SOM, and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate, which determines how the weight of each node will change. 
$\tau$ is the number of total iterations.
Values for the number of neurons is arbitrary, and for various toolboxes values for one or all of $L_0$,$L(t)$, $R(t)$  and $r^t_{BMU}$ are arbitrary.
\cite{Geach12} demonstrate the algorithm of the SOM in more detail.

@@

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T} show the results of the training networks.
            As in to Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances between weight of neurons.
            As mentioned in Section~\ref{sec: method_somz}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                  \includegraphics[width=0.8\textwidth]{../image_paper2/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                    \centering \includegraphics[width=0.8\textwidth]{../image_paper2/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption[Results of training network in $1\times2$~grid]{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

            In the upper map in Fig.~\ref{fig: 1by2T}, the dark colour between two neurons indicates that the relative distance between these two groups are relatively high, and these two groups are distinguishable groups.
            In the lower part of Fig.~\ref{fig: 1by2T}, we see that the templates are divided into two groups of 5 and 7.
            Although we know from \citetalias{Kinney96} and \citetalias{Hossein12} that 6 of the templates are quiescent and the other 6 are starbursts, the SOM results show 5 of the galaxies in one group and the other 7 in the second group.
            In this method, the Sc template has been categorized as starburst due to the relatively higher disk stellar population relative to that of the bulge. 
            According to \citetalias{Kinney96}, Sc galaxies are considered to be late Hubble type galaxies which have a flattter spectra compare to other quiescent galaxies.

@@

            Comparing Fig.~\ref{fig: 1by2T} to Fig.~\ref{fig: 1by3T} shows that the starburst templates are divided into two groups. 
            Based on the colours between these two groups, we conclude that they are very similar; both groups are starbursting and have strong emission lines.
            On the other hand, the SB5 and SB6 templates have the highest internal extinctions; this causes the spectra to become flatter at shorter wavelengths. 
            The flatter UV spectra makes these two templates more similar to quiescent galaxies than to other starbursts.
            Therefore, in the networks, SB5 and SB6 types are to be grouped close to the quiescent galaxy templates.

@@

            In Fig.~\ref{fig: 1by22T}, we can see twelve different groups:
            5 groups in the left-side neurons are separated from 7 groups on the right side of the map with two dark-grey colours between them.
            This shows that even though the galaxies are clustered in twelve groups, there still remain only two main distinct groups.
            The closest occupied neuron in the starburst side of the SOM, belongs to the SB6 type. 
            This template has the most extinction and its spectra has more similarity to quiescent galaxies than other starburst types.

@@

        \subsubsection{Classifying the Galaxy Sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Section~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have spectral types similar to those of the quiescent galaxies and the spectral types of the rest are similar to the starburst galaxies.
            The lower left panel shows the median spectra of the 80 galaxies that were classified as quiescent. 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the spectra clearly show the 4000\AA~break, one of the signatures of quiescent galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar spectral types to Sa galaxies, but with stronger emission lines.
            The median spectra in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption[Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~networks]{Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~network trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 spectra are classified into the quiescent group while the 62 of the 142 spectra are categorized as starburst galaxies. Lower panel: median spectra of the galaxies in each group, early-type on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

            Since in this network galaxies were forced to be divided into a maximum of two groups, the strongest feature in a galaxy's spectra predominantly decides which group the galaxy belongs to.
            Galaxies with a weak 4000\AA~break but strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000\AA~break are categorized as quiescent.
            Increasing the size of the SOM helps solve the problem of the galaxies which have features that are common in both groups.

@@

            Comparing the upper panel of Fig.~\ref{fig: 1by22V} with the lower panel of Fig.~\ref{fig: 1by22T} shows that the occupied neurons are not necessarily the same.
            If a cluster from the \citetalias{Hossein12} sample fills the same neuron as a \citetalias{Kinney96} template, we can conclude that spectra of galaxies in the cluster are very similar to those of the template.
            Otherwise, we can conclude that the spectra of galaxies in the cluster are (dis)similar to two of the \citetalias{Kinney96} templates.
            For the latter case, the colours establish which template has greater similarity to the \citetalias{Hossein12} cluster.

            The first two neurons in the upper plot in Fig.~\ref{fig: 1by22V} are empty, while these neurons were occupied by galaxies B and E in the trained network.
            We therefore conclude that there are no galaxies in the \citetalias{Hossein12} sample with spectral type similar to that of type B or E.
            There are 3 galaxies in the third neuron, similar to the S0 type. 
            19 of the galaxies are jn the fifth neurons and similar to the Sa type, while the 6 galaxies in the fourth neurons have spectral type similar to both S0 and Sa galaxies.
            There are 8 Sb type galaxies and 19 galaxies with spectral types similar to both Sa and Sb type galaxies.
            The following two neurons (eighth and ninth neurons from the left) are on the edge of the quiescent and starburst galaxies.
            In the upper panel in Fig.~\ref{fig: 1by22T}, the colour between these two neurons is black, which shows that these two are very different.
            Therefore, 24 galaxies in the eighth neurons have similar spectra to Sb type galaxies, and the spectra of the 5 galaxies in the ninth neuron are similar to the SB6 galaxies.
            The far right neurons in the network belong to galaxies with spectra comparable to types SB1 and SB2.

@@

            With other methods (e.g. $\chi^2$ fitting), trained neural network fitting of spectra is limited to models and the results are always reported with uncertainties.
            However, with SOM, we have the freedom to categorize galaxies in intermediate groups.
            Therefore, if the spectra of galaxies do not match completely with any of the spectra templates in Fig.~\ref{fig: k96}, they can be categorized as a spectra with similarity to two or more of the groups.

@@

        Fig.~\ref{fig: props} shows the median of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.%, and vertical axis is one of the aforementioned properties.
%        
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from quiescent galaxies to starburst. 
       Separating galaxies based on spectral types also leads to a separation in properties that derived (via {\em CIGALE}) from the spectra, as expected since the spectra types are also based on {\em CIGALE} fitting.

@@

Two-dimensional networks can be used to find the relations of the neurons and classifications in more details.
The sample galaxies have more freedom to be categorized, and only 23 of them occupies exactly the same neurons as \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
Additionally, by using two-dimensional networks we showed that we do not have nay outlier in the sample galaxies.

>>>>

\chapter[Classifying High-Redshift Galaxy Spectra]{Clustering galaxy spectra at $0.5<z<1$: an unsupervised approach}
\label{ch: paper2}
\defcitealias{Hossein12}{T12}
\defcitealias{Kinney96}{K96}
\defcitealias{Noll09}{N09}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction}
\label{sec: intro_somz}
%General information about SEDs
Nearly all information we can obtain from a galaxy is encapsulated in the light it emits; every observable phenomenon in a galaxy leaves a footprint on its spectral energy distribution (SED).
We can determine some physical property of a galaxy by properly modeling various features observed in its SED, and the general shape of the SED can be used as an identifier of the morphological type.
Considering the main features of SEDs, galaxies can be categorized into two main groups: elliptical or spiral. %PB: quiescent or SB?
Each group has its own characteristic features and can in turn be divided into many sub-branches.

@@

Advances in techniques and detectors have resulted in more detailed observations of SEDs and also made classification of galaxies more complex.
Since no two galaxies, even with the same morphology, have exactly the same properties, classifying SEDs of galaxies using templates is very challenging.
To overcome this challenge, many fitting methods have been developed and used to find template matches for SEDs, with $\chi^2$ minimization being the most commonly used. 
Artificial neural networks, K-means clustering, and principal component analysis are some other methods used to cluster and classify spectral types of galaxies \citep[e.g.][]{Allen13,Ordov14,Shi15}.

@@

%SOM
A Kohonen self-organizing map (also called self-organizing map, or SOM) is an unsupervised neural network for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
%SOM in Astronomy
The utilization of the SOM in astronomy dates back to the 1990s, with \citet[][]{Odewahn92}, \citet[][]{Hernandez94}, and \citet[][]{Murtagh95} among the first to use SOMs in their studies.
\citet{Geach12} used COSMOS data to demonstrate two of the main applications of SOMs: object classification and clustering, and photometric redshift estimation. The latter has been the subject of many other studies \citep[e.g.][]{Kind14a}.
From classifying quasars' spectra to star/galaxy classifications, from gamma-ray burst clustering to classification of light curves, this method has proved to be useful in various fields of astronomy \citep[e.g.][]{Maehoenen95, Miller96, Andreon00, Balastegui01, Rajaniemi02, Brett04, Scaringi09}.

%SOM in Astronomy continued.
Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can distinguish an astronomical object from instrumental errors, and then classify the object based on its spectrum.

%What \citetalias{Hossein12} did
~\citet[][hereafter \citetalias{Hossein12}]{Hossein12} classified spectral types of a sample of 142 galaxies using a supervised (Multi-layer feed-forward) neural network method, based on the spectral templates presented by \citetalias{Kinney96}.
With the supervised method, they could only classify spectral types of 105 out of the 142 galaxies.
The spectra of the 37 galaxies from their samples could not be matched with any of those in the \citetalias{Kinney96} templates. 
In order to classify the remaining 37 galaxies, they combined spectra of the \citetalias{Kinney96} templates.
\citetalias{Hossein12} also showed that there are tight correlations between physical properties of galaxies, and these correlations might be different for each type of galaxy.

@@

\section{Data}
\label{sec: data_highZ}
We use spectral templates from \citetalias{Kinney96} to train neural networks.
To test the trained networks, we use SED and physical properties of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12}.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in spectral clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney Spectral Model}
     \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{../image_paper2/k96.jpg}
        \caption[\citet{Kinney96} spectral templates for 12 types of galaxies]{\citet{Kinney96} spectral templates for 12 types of galaxies from \citet{Hossein12} paper (Fig. 1 in \citet{Hossein12}). The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and SC show spectra that belong to quiescent galaxies. Starburst galaxy spectra are indicated with SB 1 to 6. Higher numbers represent more intrinsic extinction.}
        \label{fig: k96}
    \end{figure}

    \citetalias{Kinney96} used ultraviolet-optical spectra of 70 star-forming and quiescent nearby galaxies to produce a set of templates that contained 12 types of spectral templates.
    These templates have been widely used in many studies to determine morphological type of galaxies or properties of specific types of galaxies~\citep[e.g.][]{Shakouri16, Paiano16, Laporte16, Holden16}.
    \citetalias{Kinney96} stated that these templates can also be used to classify the spectra of high-redshift galaxies.

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Method
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Method}
\label{sec: method_somz}
 \subsection{Self Organizing Maps}
 \label{sec: som_SOMZ}
  The SOM is a clustering method which reduces the dimensionality of data, while preserving topological features~\citep{Kohonen98}. 
 The results of the SOM are shown with a map of neurons, which their numbers are set by the user.
 Each neuron has a fixed position in the map and may contain one or more samples from the input data in a n-dimensional space,~\boldit{V} $\in \Re^n$.
 A weight vector,~\boldit{W} $\in \Re^n$, with the same dimension as the input data, is associated with each node and will be varied during the training process.
 The process of creating an SOM happens over a series of $N$ iterations.
 In each iteration, the algorithm calculates the Euclidean distance for each node $j$ as $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and finds a neuron with $D_{j_,\mathrm{min}}$. 
 This neuron is the winner node and is calling Best Matching Unit (BMU). 
 In each iteration (t), the weight vectors W(t) in the neighbourhood of the BMU ($r^t_{BMU}$) will change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
  \begin{equation}
            \label{equ: weight adj}
            W(t+1)=W(t)+L(t) \times R(t) \times(V(t)-W(t))
 \end{equation}
where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor with $L_0$ is the initial value of learning, which prevents the divergence of the SOM, and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate, which determines how the weight of each node will change. 
$\tau$ is the number of total iterations.
Values for the number of neurons is arbitrary, and for various toolboxes values for one or all of $L_0$,$L(t)$, $R(t)$  and $r^t_{BMU}$ are arbitrary.
\cite{Geach12} discussed the SOM algorithm in more detail.

@@

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T} show the results of the training networks.
            As in Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances between weight of neurons.
            As mentioned in Section~\ref{sec: method_somz}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                  \includegraphics[width=0.8\textwidth]{../image_paper2/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                    \centering \includegraphics[width=0.8\textwidth]{../image_paper2/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption[Results of training network in $1\times2$~grid]{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

            In the upper map in Fig.~\ref{fig: 1by2T}, the dark colour between two neurons indicates that the relative distance between these two groups are relatively high, and these two groups are distinguishable groups.
            In the lower part of Fig.~\ref{fig: 1by2T}, we see that the templates are divided into two groups of 5 and 7.
            Although we know from \citetalias{Kinney96} and \citetalias{Hossein12} that 6 of the templates are quiescent and the other 6 are starbursts, the SOM results show 5 of the galaxies in one group and the other 7 in the second group.
            In this method, the Sc template has been categorized as starburst due to the relatively higher disk stellar population relative to that of the bulge. 
            According to \citetalias{Kinney96}, Sc galaxies are considered to be late Hubble type galaxies which have  flattter spectra compare to other quiescent galaxies.

@@

            Comparing Fig.~\ref{fig: 1by2T} to Fig.~\ref{fig: 1by3T} shows that the starburst templates are divided into two groups. 
            Based on the colours between these two groups, we conclude that they are very similar; both groups are starbursting and have strong emission lines.
            On the other hand, the SB5 and SB6 templates have the highest internal extinctions; this causes the spectra to become flatter at shorter wavelengths. 
            The flatter UV spectra make these two templates more similar to quiescent galaxies than to other starbursts.
            Therefore, in the networks, SB5 and SB6 types are grouped close to the quiescent galaxy templates.

@@

            In Fig.~\ref{fig: 1by22T}, we can see twelve different groups:
            5 groups in the left-side neurons are separated from 7 groups on the right side of the map with two dark-grey colours between them.
            This shows that even though the galaxies are clustered in twelve groups, there still remain only two main distinct groups.
            The closest occupied neuron in the starburst side of the SOM, belongs to the SB6 type. 
            This template has the most extinction and its spectrum has more similarity to quiescent galaxies than other starburst types.

@@

        \subsubsection{Classifying the Galaxy Sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Section~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have spectral types similar to those of the quiescent galaxies and the spectral types of the rest are similar to the starburst galaxies.
            The lower left panel shows the median spectrum of the 80 galaxies that were classified as quiescent. 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the spectra clearly show the 4000\AA~break, one of the signatures of quiescent galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar spectral types to Sa galaxies, but with stronger emission lines.
            The median spectrum in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption[Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~networks]{Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~network trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 spectra are classified into the quiescent group while the 62 of the 142 spectra are categorized as starburst galaxies. Lower panel: median spectra of the galaxies in each group (fluxes are normalized by dividing spectra by flux at 4000\AA), early-type on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

            Since in this network galaxies were forced to be divided into a maximum of two groups, the strongest feature in a galaxy's spectrum predominantly decides which group the galaxy belongs to.
            Galaxies with a weak 4000\AA~break but strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000\AA~break are categorized as quiescent.
            Increasing the size of the SOM helps solve the problem of the galaxies which have features that are common in both groups.

@@

            Comparing the upper panel of Fig.~\ref{fig: 1by22V} with the lower panel of Fig.~\ref{fig: 1by22T} shows that the occupied neurons are not necessarily the same.
            If a cluster from the \citetalias{Hossein12} sample fills the same neuron as a \citetalias{Kinney96} template, we can conclude that the spectra of galaxies in the cluster are very similar to those of the template.
            Otherwise, we can conclude that the spectra of galaxies in the cluster are (dis)similar to two of the \citetalias{Kinney96} templates.
            For the latter case, the colours establish which template has greater similarity to the \citetalias{Hossein12} cluster.

            The first two neurons in the upper plot in Fig.~\ref{fig: 1by22V} are empty, while these neurons were occupied by galaxies B and E in the trained network.
            We therefore conclude that there are no galaxies in the \citetalias{Hossein12} sample with spectral type similar to that of type B or E.
            There are 3 galaxies in the third neuron, similar to the S0 type. 
            19 of the galaxies are in the fifth neuron and similar to the Sa type, while the 6 galaxies in the fourth neuron have spectral type similar to both S0 and Sa galaxies.
            There are 8 Sb type galaxies and 19 galaxies with spectral types similar to both Sa and Sb type galaxies.
            The following two neurons (eighth and ninth neurons from the left) are on the edge of the quiescent and starburst galaxies.
            In the upper panel in Fig.~\ref{fig: 1by22T}, the colour between these two neurons is black, which shows that these two are very different.
            Therefore, 24 galaxies in the eighth neuron have similar spectra to Sb type galaxies, and the spectra of the 5 galaxies in the ninth neuron are similar to the SB6 galaxies.
            The far right neurons in the network belong to galaxies with spectra comparable to types SB1 and SB2.

@@

            With other methods (e.g. $\chi^2$ fitting), trained neural network fitting of spectra is limited to models and the results are always reported with uncertainties.
            However, with SOM, we have the freedom to categorize galaxies in intermediate groups.
            Therefore, if the spectra of galaxies do not match completely with any of the spectral templates in Fig.~\ref{fig: k96}, they can be categorized as a spectrum with similarity to two or more of the groups.

@@

        Fig.~\ref{fig: props} shows the median of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.%, and vertical axis is one of the aforementioned properties.
%        
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from quiescent galaxies to starburst. 
       Separating galaxies based on spectral types also leads to a separation in properties derived (via {\em CIGALE}) from the spectra, as expected since the spectral types are also based on {\em CIGALE} fitting.

@@

Two-dimensional networks can be used to find the relations of the neurons and classifications in more details.
The sample galaxies have more freedom to be categorized, and only 23 of them occupy exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
Additionally, by using two-dimensional networks we showed that we do not have any outliers in the sample galaxies.
